Rising from the Ashes II: Can Bar Formation Lead to Abundance Bimodalities?

### 2024-09-26 paper/ms.tex
Add new figure

\documentclass[linenumbers, twocolumn]{aastex631}

@@

\title{Rising from the Ashes II: Did the Birth of the Milky Way's Bar Give Birth to the Birth of the Abundance Bimodality}

@@

\author{et al}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{fig/392276_SFH_AGN_M200_zoomed.pdf}
  \caption{\textbf{At the transition from the high- to low-$\alpha$ sequence, a bar forms and strengthens, AGN activity is enhanced, and star formation is supressed.} The evolutionary history of our SoI near the transition between the high- and low-$\alpha$ sequence, which is indicated as a vertical dashed line at an age of $10.6\,\Gyr$ ($t\sim3.2\,\Gyr$). The blue line indicates the SFR, which drops from a peak of $\sim54\,\Msunyr$ to $\sim3.6\,\Msunyr$. The minimum is very close to the transition. The drop in SFR is coincident with a rise in the AGN accretion rate (orange) as well as a rise in the strength of the bar (red).}
  \label{fig:history_zoom}
\end{figure}

We show the evolution of the SoI in greater detail at the transition between the high- and low-$\alpha$ sequences in Figure~\ref{fig:history_zoom}. As in Figure~\ref{fig:history}, we indicate the SFR (blue) and black hole accretion rate (orange), but we also include the strength of the bar, taken to be the maximum value of $A_2$ in the bar region, in red. From $\sim2.8\,\Gyr$ to $\sim3.2\,\Gyr$, the SFR drops from $\sim54$ to $3.6\,\Msunyr$. The minimum in the SFR is nearly coincident with the transition line.

We can see more clearly here that the BH accretion rate (red), in units of the Eddington accretion rate, increases from a value of $\sim0.08$ at $\sim2.8\,\Gyr$ to a maximum of $\sim0.3$ at $\sim3.1\,\Gyr$, right before the transition. This rise in the AGN accretion rate is preceded by a strengthening of the bar, which rises from an $A_2$ of $\sim0.1$ at $t\sim2.4\,\Gyr$ to an $A_2$ of $\sim0.5$ at $\sim3.1\,\Gyr$, the time of the peak BH accretion rate.

\subsection{Images}\label{ssec:images}
\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/images_star.pdf}
  \caption{Surface density projections of star particles in our SoI at the transition between the high- and low-$\alpha$ sequences. Each panel indicates subsequent snapshots ranging from snapshot 23 ($z\sim3.5$) to snapshot 36 ($z\sim1.7$), as well as snapshot 99 ($z=0$) in the lower right. Several key numbers are shown in the corners of each panel, clockwise from the top left: SFR, time, bar strength ($A_{2,\textrm{max}}$), and BH accretion rate. A guide is given in the top left subpanel. At the center-bottom of each panel, we indicate panels which occur before the transition ($\sim3.2\,\Gyr$) as high-$\alpha$ and panels which occur after as low-$\alpha$.}
  \label{fig:images_star}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/images_gas.pdf}
  \caption{As in Figure~\ref{fig:images_star}, but showing the surface density projections of gas in our SoI at the transition between the high- and low-$\alpha$ sequences. Each panel indicates subsequent snapshots ranging from snapshot 23 ($z\sim3.5$) to snapshot 36 ($z\sim1.7$), as well as snapshot 99 ($z=0$) in the lower right. Several key numbers are shown in the corners of each panel, clockwise from the top left: SFR, time, gas fraction, and BH accretion rate. A guide is given in the top left subpanel. At the center-bottom of each panel, we indicate panels which occur before the transition ($\sim3.2\,\Gyr$) as high-$\alpha$ and panels which occur after as low-$\alpha$.}
  \label{fig:images_gas}
\end{figure*}

In order to better conceptualize the evolution of the SoI at the high- to low-$\alpha$ transition, we show surface density projections of the stellar (Figure~\ref{fig:images_star}) and gaseous (Figure~\ref{fig:images_gas}) phases of the progenitors of our SoI at various redshifts. In addition to the final, $z=0$, snapshot in the lower right, we show panels for snapshots 23 to 36 in which $z$ ranges from $\sim3.5$ to $\sim1.7$ and the cosmic time from $\sim1.81$ to $\sim3.74\,\Gyr$. Projections are made on a $256\times256$ grid of a $25\,\kpc$ by $25\,\kpc$ square. All star particles/gas cells within ten times the stellar half-mass radius are used, and a simple nearest neighbor assignment of mass to each cell is used. Colors range logarithmically from $5\times10^{-4}$ to $0.5$ in native units ($10^{10} h M_{\odot} \ckpc^{-2}$), and only resolution elements within 10 times the stellar half-mass radius of the center of the subhalo are used.

We also show four relevant statistics in each panel of Figures~\ref{fig:images_star} and \ref{fig:images_gas}. For Figure~\ref{fig:images_star}, these are, clockwise from the upper left, the SFR, cosmic time, bar strength, and BH accretion rate. They are the same for Figure~\ref{fig:images_gas}, but bar strength is replaced with gas fraction (within twice the stellar half-mass radius). The bar strength is missing from several snapshots because the \citet{2022MNRAS.515.1524Z} catalog excludes certain galaxies from being considered barred if they do not pass certain quality cuts. These can come from, e.g., flybys or violent buckling events.

A description of the events portrayed in each Figure is given in Section~\ref{ssec:cause_qui}.

>>>>

\documentclass[twocolumn]{aastex631}

@@

\title{Rising from the Ashes II: Can Bar Formation Lead to Abundance Bimodalities?}

@@

\author{Vadim Semenov}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian, Cambridge, MA, USA}

\author{Lars Hernquist}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian, Cambridge, MA, USA}

\author{Vedant Chandra}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian, Cambridge, MA, USA}

\author{Charlie Conroy}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian, Cambridge, MA, USA}

@@

\subsection{Sequence of Events}\label{ssec:sequence_of_events}
\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/alyssa.pdf}
  \caption{Surface density projections of gas (top row) and star particles (middle row) in our SoI across snapshots at different times during the high- to low-$\alpha$ transition. Below the projections is a plot showing the SFR, BH accretion rate (in units of $\dot{M_{\textrm{BH}}}/\dot{M_{\textrm{Edd}}}$), and bar strength ($A_2/A_0$ for $R<2$ kpc). Time ranges from $\sim2.4\,\textrm{Gyr}$ to $\sim3.6\,\textrm{Gyr}$, corresponding to redshifts from $z\sim2.7$ to $z\sim1.8$. A sequence of events in which the bar strengthens, BH accretion increases, and SFR declines is seen, and is described more fully in the text.}
  \label{fig:seq}
\end{figure*}

Figure~\ref{fig:seq} presents surface density projections of gas and stars at different stages of the high- to low-$\alpha$ transition. The upper panels show gas density, while the middle panels display stellar density. Time progresses from $\sim2.4$ to $\sim3.6\,\textrm{Gyr}$, corresponding to redshifts ranging from $z\sim2.7$ to $z\sim1.8$.

This Figure shows the following sequence of events:
\begin{enumerate}
    \item Bar forms: A steady increase in the bar strength, as indicated by $A_2/A_0$ for star particles with $R<2\,\kpc$, from $\sim0.05$ to $0.4$ starting around $2.8\,\textrm{Gyr}$. This rise is accompanied by the appearance of elongated features in the gas and stars consistent with a bar.
    \item BH accretion increases: Following the increase in bar strength by about a snapshot ($\sim150\,\Myr$ here), the black hole accretion rate ($\dot{M_{\textrm{BH}}}/\dot{M_{\textrm{Edd}}}$) shows a significant spike, rising from a minimum of $\sim0.08$ at $t=2.84\,\Gyr$ to a maximum of $\sim0.29$ at $t=3.13\,\Gyr$.
    \item SFR declines: After the increase in black hole accretion, the star formation rate (SFR) begins to decline. The decrease in SFR is gradual and continues as the black hole accretion rate remains elevated. By approximately $3.2\,\textrm{Gyr}$, the SFR reaches a more quiescent state, where it remains relatively low.
\end{enumerate}
